A Plasma Torus Around a Young Low-Mass Star

### 2025-03-14 paper/ms.tex
update

The material's composition -- either pure plasma, or else a dusty
plasma -- can be clarified by time-series optical and infrared
spectrophotometry.  While observations of CPVs in the optical have
previously suggested chromaticity consistent with dust \cite{Tanimoto2020,Gunther2022,Koen2023},
a gray opacity source such as electron scattering in a plasma
transiting over starspots could also produce chromatic features
\cite{Rackham2018}.  The composition and size distribution of any dust
that is present could be most easily resolved by measuring the
extinction curve for one or more CPVs from 1-10\,$\mu$m.  A
dust composition similar to debris from rocky bodies seen around white
dwarfs \cite{Reach2009} would indicate a rocky-body origin.  A 
composition closer to the ISM would be indicative of condensed dust in
an M dwarf wind, similar to that formed in the environments of more
evolved stars \cite{Marigo2008}.

@@

{\it Viewing Orientation}---We fitted the rotational broadening of the
K~I (7700\,\AA) resonance line using the broadening kernel suggested by
\cite{Gray2008}; taking the mean and standard deviation of the
resulting value over all epochs yielded $v_{\rm eq} \sin
i$=138$\pm$8\,\kms, consistent with the visual line broadening $\Delta
\lambda$$\approx$3\,\AA.  The star's equatorial velocity $v_{\rm eq}$
based on its apparent size and rotation period is 130$\pm$4\,\kms.
While this suggests that the viewing orientation could be nearly
edge-on, the formal constraint is rather weak, with $i$$>$63$^\circ$ at
2$\sigma$ (2.5$^{\rm th}$ percentile of the inclination posterior).
% from calc_max_amplitude_sinusoid_rv_timeseries.py
% from K_to_msini.py, assuming Mstar = 0.18

@@

{\it Effective temperature, radius, mass, and spectral type}
We adopted the star's effective temperature and radius measured using
the spectral energy distribution fitting procedure described by
\cite{Bouma2024}.  To summarize, this approach used
\texttt{astroARIADNE} \cite{Vines2022} with the BT-Settl stellar
atmosphere models \cite{Allard2012}, assuming the \cite{Asplund2009}
solar abundances and the \cite{Barber2006} water line lists.  This
approach fitted for the stellar effective temperature, radius,
reddening, surface gravity, metallicity, and distance by comparing the
measured broadband magnitudes against pre-computed model grids.
Specifically, we used the broadband magnitudes from Gaia DR2, APASS,
2MASS, SDSS, and WISE $W1$ and $W2$.  This method has the most
constraining power for the star's effective temperature (2972 $\pm$
40\,K) and radius (0.42 $\pm$ 0.02\,$R_\odot$); the surface gravity,
metallicity, and reddening are typically only weakly constrained.  We
determined the star's spectral type to be M5.5Ve by visually comparing
against the photometric standards tabulated by \cite{Bochanski2007},
which agrees with the effective temperature from the SED fitting
\cite{Pecaut2013}.

@@

{\it SED: No evidence for disk}---Figure~\ref{fig:sed} shows the 
...

>>>>

The material's composition -- either pure plasma, or else a dusty plasma
-- can be clarified by time-series optical and infrared
spectrophotometry.  While observations of CPVs in the optical have
previously suggested chromaticity consistent with dust
\cite{Tanimoto2020,Gunther2022,Koen2023}, a gray opacity source such as
electron scattering in a plasma transiting over starspots might also
produce chromatic features \cite{Rackham2018}.  The composition and size
distribution of any dust that is present would be most easily resolved
by measuring the extinction curve for one or more CPVs from
1-10\,$\mu$m.  Composition and size distributions similar to debris from
rocky bodies seen around white dwarfs \cite{Reach2009} would indicate a
rocky-body origin.  Compositions and sizes to the ISM would be
indicative of condensed dust in an M dwarf wind, similar to that formed
in the environments of more evolved stars \cite{Marigo2008}.

@@

{\it Viewing Orientation}---We fitted the rotational broadening of the
K~I (7700\,\AA) resonance line using the broadening kernel suggested by
\cite{Gray2008}; taking the mean and standard deviation of the resulting
value over all epochs yielded $v_{\rm eq} \sin i$=138$\pm$8\,\kms,
consistent with the visual line broadening $\Delta
\lambda$$\approx$3\,\AA.  The star's equatorial velocity $v_{\rm eq}$
based on its apparent size and rotation period is 130$\pm$4\,\kms.
While this suggests that the viewing orientation could be nearly
edge-on, the formal constraint is rather weak, with $i$$>$63$^\circ$ at
2$\sigma$ (2.5$^{\rm th}$ percentile of the inclination posterior).

% from calc_max_amplitude_sinusoid_rv_timeseries.py
% from K_to_msini.py, assuming Mstar = 0.18

@@

\begin{figure}[!t]
  \centering
  \includegraphics[width=0.7\textwidth]{figures/sf4.pdf}
  \caption{
    Spectral energy distribution of broadband photometric magnitudes
    plotted over the best-fit BT-Settl stellar atmosphere model
    \cite{Allard2012}.  This plot was made using \texttt{astroARIADNE}
    \cite{Vines2022}.  The photometry extends from the Gaia DR3 blue
    passband to WISE W3;  the redder (22 $\mu$m) W4 passband did not yield
    a confident detection.  This fit yields the star's temperature and
    size.  The lack of excess infrared flux relative to the photospheric
    model sets an upper limit on emission from any circumstellar dust.
    }
  \label{fig:sed}
\end{figure}

{\it Effective temperature, radius, mass, and spectral type} We adopted
the star's effective temperature and radius measured using the spectral
energy distribution (SED) fitting procedure described by
\cite{Bouma2024}.  To summarize, this approach used
\texttt{astroARIADNE} \cite{Vines2022} with the BT-Settl stellar
atmosphere models \cite{Allard2012}, assuming the \cite{Asplund2009}
solar abundances and the \cite{Barber2006} water line lists.  This
approach fitted for the stellar effective temperature, radius,
reddening, surface gravity, metallicity, and distance by comparing the
measured broadband magnitudes against pre-computed model grids.
Specifically, we used the broadband magnitudes from Gaia DR2, APASS,
2MASS, SDSS, and WISE $W1$ and $W2$.  The resulting best-fit SED is
shown in Figure~\ref{fig:sed}.  This method has the most constraining
power for the star's effective temperature (2972 $\pm$ 40\,K) and radius
(0.42 $\pm$ 0.02\,$R_\odot$); the surface gravity, metallicity, and
reddening are typically only weakly constrained.  We determined the
star's spectral type to be M5.5Ve by visually comparing against the
photometric standards tabulated by \cite{Bochanski2007}, which agrees
with the effective temperature from the SED fitting \cite{Pecaut2013}.

@@

{\it Upper and Low Bounds on Dust}---An upper limit on the amount of hot
dust follows from the lack of an infrared excess.  A lower limit follows
from the assumption that most of the broadband optical depth comes from 
dust, rather than some other source.
Figure~\ref{fig:sed} shows the SED; ...
%TODO TODO LEFT OFF HERE
...
